---
title: "Wave energy production"
author: "*Raynaldi Pratama*"
date: "15/03/2021"
output: 
  rmdformats::downcute:
    self_contained: TRUE
    thumbnails: TRUE
    lightbox: TRUE
    gallery: TRUE
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Acknowledgement:** This portfolio is a re-worked exercise, part of the Ocean Wave Energy and Offshore Wind Energy Assessment course, conducted during Spring 2019 at *Universidad del País Vasco/Euskal Herriko Unibertsitatea* (UPV/EHU), Spain. This document aims to visualize the energy production data from the 2015 records of the existing wave energy converter (WEC) facility situated in Mutriku.

**Library setup:**

```{r warning=FALSE, message=FALSE}
# Working with table
library(kableExtra)

# World map data from Natural Earth
library(rnaturalearth)
library(rnaturalearthdata)

# Data visualization
library(ggplot2)
library(ggspatial)

# Manipulating and analysing data
library(tidyverse)
```

## Mutriku WEC location

The subject of analysis, [Mutriku wave farm](https://tethys.pnnl.gov/project-sites/mutriku-wave-power-plant), is situated $\pm$ 65.6 km to the SE from Bilbao-Vizcaya buoy location. Mutriku wave farm adopts the oscillating water column (OWC) technology in which 16 wells-type turbines (18.5 kW each) are installed inside the breakwater structure. For the turbines control purpose, the plant is equipped with the SCADA system where the essential historical data may be acquired.

### Load the spatial data

```{r}
spdf_world <- ne_countries(scale = "medium", returnclass = "sf")

sldf_world <- ne_coastline(scale = "medium", returnclass = "sf")
```

### Plot the corresponding data

```{r fig.align='center', message=FALSE}
theme_set(theme_bw())

ggplot() +
  geom_sf(data = spdf_world) +
  geom_sf(data = sldf_world, size = 0.75) +
  annotation_scale(width_hint = 0.5, location = "bl") +
  annotation_north_arrow(
    location = "tl",
    height = unit(1.5, "cm"),
    width = unit(1.5, "cm"),
    pad_x = unit(0.25, "cm"),
    pad_y = unit(0.25, "cm"),
    style = north_arrow_fancy_orienteering
  ) +
  annotate(
    geom = "text",
    x = -6,
    y = 46,
    label = "Bay of Biscay",
    fontface = "italic",
    color = "darkblue",
    size = 4
  ) +
  annotate(
    geom = "point",
    x = -2.3781,
    y = 43.312,
    colour = "blue",
    pch = 17,
    size = 5
  ) +
  annotate(
    geom = "text",
    x = -2.35,
    y = 43.65,
    label = "Mutriku WEC",
    size = 3
  ) +
  coord_sf(xlim = c(-12, 2), ylim = c(40, 50), expand = FALSE) +
  theme(
    panel.grid.major = element_line(
      color = gray(0.5), linetype = "dashed", size = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Mutriku WEC location")
```

## Generated wave energy

### Load the data

```{r}
energy <- read.table("KWH_2015.txt", header = TRUE)

kbl(energy) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")

summary(energy)
```

### Energy injected to the grid

```{r}
energy_Max <- max(energy$cumulated_kwh, na.rm = TRUE) # Maximum energy generation in kWh

energy_Min <- min(energy$cumulated_kwh, na.rm = TRUE) # Minimum energy generation in kWh

energy_Inj <- energy_Max - energy_Min

energy_Inj # Total energy injected (in kWh) to the grid in 2015
```

### WEC working hours

```{r}
workinghrs <- sum(!is.na(energy$kwh))

workinghrs # Annual working hours in 2015

percentage <- 100 * workinghrs / 8760

round(percentage, 2) # Annual working hours (in %) in 2015
```

## Monthly wave energy production

### Average hourly EP

```{r}
energy_Avg <- energy_Inj / workinghrs

energy_Avg # in kW in an hour
```
### Box plot

```{r fig.align='center', warning=FALSE}
ggplot(energy, aes(x = factor(month), y = kwh, fill = factor(month))) +
  geom_boxplot(notch = TRUE, outlier.alpha = 0.1, varwidth = TRUE) +
  geom_hline(
    yintercept = energy_Avg,
    linetype = "dashed",
    color = "blue",
    size = 0.75
  ) +
  geom_text(
    x = 3,
    y = 150,
    color = "blue",
    label = paste("Average hourly EP = ",
      round(energy_Avg, 2), " kWh",
      sep = ""
    )
  ) +
  labs(
    x = "Month",
    y = "Hourly energy production (kWh)",
    title = "Monthly distribution of wave energy production"
  ) +
  scale_fill_discrete(
    name = "Month",
    labels = c(
      "January", "February", "March", "April", "May", "June",
      "July", "August", "September", "October", "November", "December"
    )
  )
```

From the figure above, it is shown that the Mutriku wave farm started the year 2015 by surpassing the hourly average power production with the value of around 50 kWh. The trend began decreasing to below 37.94 kWh in April and remained under the average value until the beginning of Fall in September. The trend, later, reached its peak in November and remained over 50 kWh by the end of the year.

## Wave energy flux

### Load the data

```{r}
buoydata <- read.table("BB_2015.txt", header = TRUE)

kbl(buoydata) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")

summary(buoydata)
```

### Calculate WEF

For irregular wave, the energy transport (wave energy flux) can be calculated using the following formula:

$$J = \frac{\rho g^2}{64\pi} T_j {H_s}^2 = 0.488 \times T_j {H_s}^2$$

Where, $J$ is wave energy flux (kW/m); $\rho$ = 1.020 ton/m^3^; $g$ = 9.81 m/s^2^; $T_j$ is energy period (sec.); $H_s$ is significant wave height (m).

```{r}
wef <- 0.488 * ((buoydata$Hm0_m)^2 * buoydata$Tm02_s)

buoydataNew <- cbind(buoydata, wef)

kbl(buoydataNew) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")

wef_Avg <- mean(buoydataNew$wef, na.rm = TRUE)

wef_Avg # kW/m
```

### Box plot

```{r fig.align='center', warning=FALSE}
# Initial box plot
ggplot(buoydataNew, aes(x = factor(MM_m), y = wef, fill = factor(MM_m))) +
  geom_boxplot(notch = TRUE, outlier.alpha = 0.1, varwidth = TRUE) +
  geom_hline(
    yintercept = wef_Avg,
    linetype = "dashed",
    color = "blue",
    size = 0.75
  ) +
  geom_text(
    x = 3,
    y = 300,
    color = "blue",
    label = paste("Average WEF = ", round(wef_Avg, 2), " kW/m", sep = "")
  ) +
  labs(
    x = "Month",
    y = "Energy transport or WEF (kW/m)",
    title = "WEF from Bilbao-Vizcaya buoy data in 2015"
  ) +
  scale_fill_discrete(
    name = "Month",
    labels = c(
      "January", "February", "March", "April", "May", "June",
      "July", "August", "September", "October", "November", "December"
    )
  )
```

The same trend, as compared to the previous figure, is represented by the historical data obtained from the buoy, which located approximately 65.6 km to the Northwest from the Mutriku site. The amount of wave energy tends to be higher starting from the middle of Winter in January to the beginning of Spring in March and remains at the lower value until October. Nevertheless, the trend begins to rise again by the end of Fall in November until December. It thus indicates that the transported energy through ocean waves from the buoy along the way to Mutriku WEC has similar characteristics, which may be useful to estimate the plant's efficiency index by comparing the energy capture from both locations. Additionally, for better visualization, the log scale may be introduced in the y-axis.

```{r fig.align='center', warning=FALSE}
# Log scale box plot
ggplot(buoydataNew, aes(x = factor(MM_m), y = log(wef + 1), fill = factor(MM_m))) +
  geom_boxplot(notch = TRUE, outlier.alpha = 0.1, varwidth = TRUE) +
  geom_text(
    x = 3,
    y = 5.5,
    color = "blue",
    label = paste("Average WEF = ", round(wef_Avg, 2), " kW/m", sep = "")
  ) +
  labs(
    x = "Month",
    y = "Wave power flux in logarithmic scale",
    title = "WEF from Bilbao-Vizcaya buoy data in 2015"
  ) +
  scale_fill_discrete(
    name = "Month",
    labels = c(
      "January", "February", "March", "April", "May", "June",
      "July", "August", "September", "October", "November", "December"
    )
  )
```

## Air pressure variation

### Load the data

```{r}
press <- read.table("PRES_2015.txt", header = TRUE)

summary(press)
```

### Average air pressure

```{r}
press_Avg <- colMeans(press[, 5:20], na.rm = TRUE)

press_Avg # Average air pressure in each turbine's chamber in Pa

press_Avg_2 <- mean(press_Avg, na.rm = TRUE)

press_Avg_2 # Average air pressure of all working turbines in Pa
```

### Box plot

```{r fig.align='center'}
boxplot(press[, 5:20],
  col = "orange",
  border = "brown",
  xlab = "Turbine(s)", ylab = "Air pressure (Pa)", xaxt = "n",
  main = "Air pressure in each turbine's chamber in 2015",
  notch = TRUE,
  varwidth = TRUE
)

axis(1, at = seq(1, 16, 1), labels = names(press[5:20]), cex.axis = 0.6)

segments(0, press_Avg_2, 17, press_Avg_2, lwd = 2, lty = 2, col = "blue")

addedText3 <- paste("Average air pressure (in one year) = ",
  round(press_Avg_2, 2), " Pa",
  sep = ""
)

text(11, 19000, addedText3, font = 1, col = "blue")
```

From either the table or the box plot, we can see no value shown in turbine 01 and 16. It is likely because both turbines were being down or not in operation for some reasons. We should keep this in mind for the rest of our analysis so that only turbine 02 until turbine 15 was in service during 2015.

## Power production of each chamber

### Load the data

```{r}
power5 <- read.table("POW_2015.txt", header = TRUE)

power5_kW <- power5[, 5:20] / 1000 # Convert the unit into kW

summary(power5_kW)
```

### Average power production

```{r}
power5_Avg <- colMeans(power5_kW, na.rm = TRUE)

power5_Avg # Average power produced of each turbine in kW

power5_Avg_2 <- mean(power5_Avg, na.rm = TRUE)

power5_Avg_2 # Average power produced of all working turbines in kW
```
### Box plot

```{r fig.align='center'}
boxplot(power5_kW,
  col = "red",
  border = "brown",
  xlab = "Turbine(s)", ylab = "Power production (kW)", xaxt = "n",
  main = "Power production of each turbine",
  notch = TRUE,
  varwidth = TRUE
)

axis(1, at = seq(1, 16, 1), labels = names(press[5:20]), cex.axis = 0.6)

segments(0, power5_Avg_2, 17, power5_Avg_2, col = "blue", lwd = 2, lty = 2)

addedtext4 <- paste("Average power production = ",
  round(power5_Avg_2, 2), " kW (per 5 min. record)",
  sep = ""
)

text(4, 15, addedtext4, font = 1, col = "blue")
```

## Average valve opening degree

### Load the data

```{r}
posValve <- read.table("POSVALV_2015.txt", header = TRUE)

summary(posValve)
```

### Valve opening degree T10

```{r fig.align='center'}
plot(press[, 14], posValve[, 14],
  ylim = c(0, 90), xlab = "Air pressure (Pa)",
  ylab = "Opening degree (degrees)", main = "Turbine 10"
)

pressValve10 <- cbind(press[, 14], posValve[, 14])

degMax10 <- pressValve10[pressValve10[, 2] > 80, ]

pressMax10 <- max(degMax10[, 1], na.rm = TRUE)

pressMax10 # Maximum air pressure comes when the valve opening degree > 80 deg in Pa

segments(pressMax10, 0, pressMax10, 90, col = "blue", lwd = 2, lty = 2)
```

### Valve opening degree T15

```{r fig.align='center'}
plot(press[, 19], posValve[, 19],
  ylim = c(0, 90), xlab = "Air pressure (Pa)",
  ylab = "Opening degree (degrees)", main = "Turbine 15"
)

pressValve15 <- cbind(press[, 19], posValve[, 19])

degMax15 <- pressValve15[pressValve15[, 2] > 80, ]

pressMax15 <- max(degMax15[, 1], na.rm = TRUE)

pressMax15 # Maximum air pressure comes when the valve opening degree > 80 deg in Pa

segments(pressMax15, 0, pressMax15, 90, col = "blue", lwd = 2, lty = 2)
```

## Plant's efficiency index (PEI)

### Calculate PEI

```{r fig.align='center'}
power5_Avg_3 <- rowMeans(power5_kW, na.rm = TRUE) # Average power produced in kW per 5 minutes

PEI <- power5_Avg_3 / buoydataNew[, 7] # buoydataNew[, 7] represents wef estimation data from buoy

PEI <- na_if(PEI, "NaN")

PEI <- PEI[PEI < 0.6] # To limit the measured PEI

max(PEI, na.rm = TRUE)

min(PEI, na.rm = TRUE)

mean(PEI, na.rm = TRUE)

hist(PEI,
  xlim = c(0, 0.6), breaks = seq(0, 1, 0.025),
  freq = TRUE, col = "lightgrey", ylim = c(0, 80),
  main = "Plant's efficiency index (PEI) distribution"
)

segments(0.255, 0, 0.255, 80, col = "blue", lwd = 2, lty = 2)

text(0.35, 75, "Average PEI = 0.255", col = "blue", font = 1)
```

### Box plot

```{r fig.align='center', warning=FALSE}
PEI_df <- data.frame(cbind(PEI, power5$month))

ggplot(PEI_df, aes(x = factor(V2), y = PEI, fill = factor(V2))) +
  geom_boxplot(outlier.alpha = 0.1, varwidth = TRUE) +
  labs(
    x = "Month",
    y = "Plant's efficiency index (PEI)",
    title = "Plant's efficiency index (PEI) of Mutriku WEC in 2015"
  ) +
  scale_fill_discrete(
    name = "Month",
    labels = c(
      "January", "February", "March", "April", "May", "June",
      "July", "August", "September", "October", "November", "December"
    )
  )
```
